\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{setspace}
\onehalfspacing

\title{The Price Elasticity of AI Demand}
\author{}
\date{}

\begin{document}
\maketitle

\section*{Question}

How elastic is consumer demand for AI subscriptions?

\section*{Setting}

Chicago levies the Personal Property Lease Transaction Tax (PPLTT) on cloud software, including AI subscriptions. No other major US city does this. The rate increased twice:

\begin{center}
\begin{tabular}{lcc}
\toprule
Date & Rate & Change \\
\midrule
Pre-2025 & 9\% & --- \\
January 2025 & 11\% & +2pp \\
January 2026 & 15\% & +4pp \\
\bottomrule
\end{tabular}
\end{center}

Anthropic passes through the tax in full. OpenAI pass-through needs verification.

\section*{Identification}

Difference-in-differences comparing AI subscription churn in Chicago versus surrounding suburbs.

\textbf{Treatment:} Chicago ZIP3s (606, 608)

\textbf{Control:} Suburban Cook County and collar counties (600--605, 607)

\textbf{Unit:} Individual cardholder $\times$ month

\textbf{Outcome:} Churn (no AI subscription transaction for $k$ consecutive months after prior subscription)

\section*{Specification}

\begin{equation}
\text{Churn}_{it} = \beta_1 (\text{Chicago}_i \times \text{Post2025}_t) + \beta_2 (\text{Chicago}_i \times \text{Post2026}_t) + \gamma_i + \delta_t + \epsilon_{it}
\end{equation}

$\gamma_i$ = cardholder fixed effects

$\delta_t$ = month fixed effects

$\beta_1$ identifies the effect of the 2pp increase; $\beta_2$ identifies the cumulative effect through the 4pp increase.

\section*{Data}

Individual-level credit and debit card panel. Can isolate transactions to OpenAI, Anthropic, and related AI merchants. ZIP3 geography links cardholders to Chicago vs.\ suburbs.

\section*{Contribution}

First causal estimate of the price elasticity of demand for AI tools.

\section*{What Could Kill This}

\begin{enumerate}
\item \textbf{Sample size.} Fewer than 500 Chicago AI subscribers pre-2025 means too little power.
\item \textbf{No pass-through.} If OpenAI absorbs the tax, only Anthropic users are treated—smaller sample.
\item \textbf{Pre-trends.} If Chicago and suburban churn rates were already diverging before January 2025, the DiD is contaminated.
\item \textbf{Billing avoidance.} If Chicago users switch billing addresses to suburbs, true treatment effect is attenuated.
\end{enumerate}

\section*{Extensions (If Core Works)}

\begin{itemize}
\item Heterogeneity by income (proxy: total card spend)
\item Heterogeneity by self-employment (proxy: business-related spending)
\item Welfare calculation using external productivity estimates
\end{itemize}

These are gravy. The elasticity is the contribution.

\end{document}
